A Data-Driven Framework for Neural Field Modelling

### 2010-11-18 ltx/first_order_framework/SupportingInformation.tex
changed to Neuroimage cls file and added appendices, still need to fix long equations and cross referencing

\section*{Spatial Cut-off Frequency of Field Basis Functions}\label{ap:FrequencyAnalysis}
This section gives the derivation for the width of the Fourier transform of  the field basis function in terms of its spatial cut-off frequency (equations~34 and~37 in the main text). We define an \textit{n}-dimensional Gaussian  centred at the origin as 
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r\right).
\end{equation}
Now applying the \textit{n}-dimensional Fourier transform yields
\begin{align}\label{eq:AppendixGaussianFT}
 \boldsymbol\Phi(\boldsymbol \nu)&=\int_{\mathbb{R}^n} {\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r} \nonumber \\
&=\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \times\mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right).
\end{align}
Evaluating the integral on the RHS of equation~\ref{eq:AppendixGaussianFT} gives 
\begin{equation}\label{eq:IntegralOfGaussian}
\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r
=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}.
\end{equation}
Substituting equation~\ref{eq:IntegralOfGaussian} into equation~\ref{eq:AppendixGaussianFT} gives the scaled Gaussian
\begin{equation}\label{eq:FTofGaussian}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation} 
where 
\begin{equation}
	\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
For $3$~dB attenuation ($1/2$ to peak power) at the cut-off frequency, $\boldsymbol\nu_c$, we need to set
\begin{equation}
 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2,
\end{equation}
since the maximum of Gaussian described in equation~S1.10 is at the origin. Now solving for $\sigma_{\nu}^2$ we get
\begin{equation}
 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_{c}^\top \boldsymbol\nu_{c}}{\ln 2 }.
\end{equation}

\section*{Convolution of Two $n$-Dimensional Gaussians}\label{ap:ConvOfGaussians}
In this section, we provide a derivation for the convolution of two $n$-dimensional Gaussian basis functions. This derivation is used in the analytic calculation of $\boldsymbol\Psi(\mathbf{r})$ and the observation matrix, $\mathbf C$. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}
The derivation proceeds by assuming the Gaussian basis functions are positioned at the origin ($\boldsymbol\mu_i=\boldsymbol\mu_j=\mathbf 0$). The convolution theorem states that the Fourier transform of the convolution of two functions is the product of their Fourier transforms. Using the derivation of the Fourier transform of a \emph{n}-dimensional Gaussian from equation~S1.10 of the previous section, we have
\begin{equation}\label{eq:FTTwoGaussianMultiplication}
\mathcal{F}\left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\pi^2\sigma_i^2\sigma_j^2\right)^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_i^2+\sigma_j^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right).
\end{equation}
where $\mathcal{F}$ denotes the Fourier transform. Multiplying and dividing by $\left(\pi\left(\sigma_i^2+\sigma_j^2\right)\right)^{\frac{n}{2}}$ equation~\ref{eq:FTTwoGaussianMultiplication} can be written as 
\begin{equation}
\mathcal{F}\left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\pi\sigma_i^2\sigma_j^2\right)^{\frac{n}{2}}\left(\frac{1}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} (\pi(\sigma_i^2+\sigma_j^2))^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_i^2+\sigma_j^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right).
\end{equation}
Taking the inverse Fourier transform, the convolution becomes
\begin{equation}\label{eq:ConvoftwoGaussian}
 \left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} \mathbf r^\top\mathbf r}\right).
\end{equation}
In the case where $\boldsymbol\mu_i\neq\boldsymbol\mu_j\neq\mathbf 0$, equation~\ref{eq:ConvoftwoGaussian} becomes
\begin{equation}\label{eq:ConvoftwoGaussianNonzeroMean}
 \left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left( \frac{\pi\sigma_i^2\sigma_j^2} {\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\mathbf r-\boldsymbol\mu)^\top(\mathbf r-\boldsymbol\mu)}\right),
\end{equation}
where $\boldsymbol\mu=\boldsymbol\mu_i+\boldsymbol\mu_j$.

@@

\section*{Observation Kernel}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Figure~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered by visually using the microscope. 
\begin{figure}[ht]		
	\begin{center}
		%trim option's parameter order: left bottom right top
	\includegraphics[scale=.1, trim = .5mm .5mm .00005mm .00005mm, clip ] {./Graph/eps/SensorAndSource.eps}
\end{center}
	\caption{Photograph through a microscope of the Utah recording array and the twisted bipolar current source.}
	\label{fig:SensorAndSource}
\end{figure}

The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used estimate the shape of the sensor kernel. To extract the amplitude the signal the data was first band-pass filtered with a centre frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{N}\sum_n{\sqrt{(x(n)-\mathcal{H}x(n))^2}}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $n$ indexes samples and $N$ is the total number of samples in the time series. Figure~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus to distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution. 
\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=1]{./Graph/eps/channel3fig.eps}
	\end{center}
	\caption{Amplitude of the envelope of the signal recorded from an electrode at the origin as the source was moved.}
	\label{fig:Ch3}
\end{figure}

>>>>

\section*{Spatial Cut-off Frequency of Field Basis Functions}\label{ap:FrequencyAnalysis}
This section gives the derivation for the width of the Fourier transform of  the field basis function in terms of its spatial cut-off frequency (equations~34 and~37 in the main text). We define an \textit{n}-dimensional Gaussian  centred at the origin as 
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r\right).
\end{equation}
Now applying the \textit{n}-dimensional Fourier transform yields
\begin{align}\label{eq:AppendixGaussianFT}
 \boldsymbol\Phi(\boldsymbol \nu)&=\int_{\mathbb{R}^n} {\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r} \nonumber \\
&=\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \times\mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right).
\end{align}
Evaluating the integral on the RHS of equation~\ref{eq:AppendixGaussianFT} gives 
\begin{equation}\label{eq:IntegralOfGaussian}
\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r
=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}.
\end{equation}
Substituting equation~\ref{eq:IntegralOfGaussian} into equation~\ref{eq:AppendixGaussianFT} gives the scaled Gaussian
\begin{equation}\label{eq:FTofGaussian}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation} 
where 
\begin{equation}
	\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
For $3$~dB attenuation ($1/2$ to peak power) at the cut-off frequency, $\boldsymbol\nu_c$, we need to set
\begin{equation}
 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2,
\end{equation}
since the maximum of the Gaussian described in equation~S1.10 is at the origin. Solving for $\sigma_{\nu}^2$, we get
\begin{equation}
 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_{c}^\top \boldsymbol\nu_{c}}{\ln 2 }.
\end{equation}

\section*{Convolution of Two $n$-Dimensional Gaussians}\label{ap:ConvOfGaussians}
In this section, we provide a derivation for the convolution of two $n$-dimensional Gaussian basis functions. This derivation is used in the analytic calculation of $\boldsymbol\Psi(\mathbf{r})$ and the observation matrix, $\mathbf C$. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}
The derivation proceeds by assuming the Gaussian basis functions are positioned at the origin ($\boldsymbol\mu_i=\boldsymbol\mu_j=\mathbf 0$). The convolution theorem states that the Fourier transform of the convolution of two functions is the product of their Fourier transforms. Using the derivation of the Fourier transform of a \emph{n}-dimensional Gaussian from equation~S1.10 of the previous section, we have
\begin{equation}\label{eq:FTTwoGaussianMultiplication}
\mathcal{F}\left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\pi^2\sigma_i^2\sigma_j^2\right)^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_i^2+\sigma_j^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right),
\end{equation}
where $\mathcal{F}$ denotes the Fourier transform. Multiplying and dividing by $\left(\pi\left(\sigma_i^2+\sigma_j^2\right)\right)^{\frac{n}{2}}$, equation~\ref{eq:FTTwoGaussianMultiplication} can be written as 
\begin{equation}
\mathcal{F}\left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\pi\sigma_i^2\sigma_j^2\right)^{\frac{n}{2}}\left(\frac{1}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} (\pi(\sigma_i^2+\sigma_j^2))^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_i^2+\sigma_j^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right).
\end{equation}
Taking the inverse Fourier transform, the convolution becomes
\begin{equation}\label{eq:ConvoftwoGaussian}
 \left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} \mathbf r^\top\mathbf r}\right).
\end{equation}
In the case where $\boldsymbol\mu_i\neq\boldsymbol\mu_j\neq\mathbf 0$, equation~\ref{eq:ConvoftwoGaussian} becomes
\begin{equation}\label{eq:ConvoftwoGaussianNonzeroMean}
 \left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left( \frac{\pi\sigma_i^2\sigma_j^2} {\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\mathbf r-\boldsymbol\mu)^\top(\mathbf r-\boldsymbol\mu)}\right),
\end{equation}
where $\boldsymbol\mu=\boldsymbol\mu_i+\boldsymbol\mu_j$.

@@

\section*{Observation Kernel}
The Gaussian shape of the observation kernel was verified experimentally. This was achieved by moving a dipole across a recording electrode submerged in saline. A photograph through a microscope of the submerged electrodes can be seen in Figure~\ref{fig:SensorAndSource}. The measurement sensor was the most central electrode of a $5\times5$ Utah array (Blackrock Microsystems) which had electrode spacing of 400~$\mu$m. The current source was a twisted wire electrode (Plastics One Inc.) with tip-tip spacing of 60~$\mu$m and tip diameter of 200~$\mu$m. The twisted wire electrode was moved in the saline solution using a 3 dimensional manipulator (Sutter Instrument Company, MPC-100) in increments of 200~$\mu$m. The twisted wire electrode and the Utah array were co-registered visually using the microscope. 
\begin{figure}[ht]		
	\begin{center}
		%trim option's parameter order: left bottom right top
	\includegraphics[scale=.1, trim = .5mm .5mm .00005mm .00005mm, clip ] {./Graph/eps/SensorAndSource.eps}
\end{center}
	\caption{Photograph through a microscope of the Utah recording array and the twisted bipolar current source.}
	\label{fig:SensorAndSource}
\end{figure}

The stimulation waveform for the current source consisted of a 40~Hz sine wave with a duration of 1~s. The recording from the central electrode in the array was used to estimate the shape of the sensor kernel. To extract the amplitude of the signal, the data was first band-pass filtered with a centre frequency of 40~Hz. The envelope of the signal was found by computing the mean instantaneous amplitude over the 1~s stimulation duration by
\begin{equation}
	a = \frac{1}{N}\sum_n{\sqrt{(x(n)-\mathcal{H}x(n))^2}}
\end{equation}
where $\mathcal{H}$ denotes the Hilbert transform operator, $n$ indexes samples and $N$ is the total number of samples in the time series. Figure~\ref{fig:Ch3} shows the mean amplitude of the recorded envelope versus the distance between the twisted wire electrode and the recording electrode. A Gaussian fitted to the amplitudes is also shown. The figure shows that a Gaussian is indeed a good model for the sensor kernel in the homogeneous solution. 
\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=1]{./Graph/eps/channel3fig.eps}
	\end{center}
	\caption{Amplitude of the envelope of the signal recorded from an electrode at the origin as the source was moved.}
	\label{fig:Ch3}
\end{figure}
